pema application

### 2023-02-15 paper/manuscript.Rmd
final revisions for publication

```{r tabr2eta, results= "asis", eval = TRUE}
tabanova <- readRDS("anova.RData")
tabanova <- tabanova[!grepl(":", tabanova$condition), ]
tabanova$Interpretation[!tabanova$Interpretation %in% c("positive", "negative", "other")] <- NA
#tabanova$condition <- gsub("alpha", "omega", tabanova$condition, fixed = T)
#tabanova$condition <- gsub("$mean{n}$", "$n$", tabanova$condition, fixed = T)
tabanova$Factor <- gsub("\\$(omega|beta|tau)", "\\$\\\\\\1", tabanova$Factor)
#tabanova$HS.vs..LASSO <- NULL
tabanova <- tabanova[, c(ncol(tabanova), 2:(ncol(tabanova)-1))]
# kbl(tabanova, caption = "Effect size of design factors and their two-way interactions on R2 the different algorithms, and of the difference between algorithms. The comparison between HS and LASSO was zero in the second decimal for all conditions. Interpretation indicates whether the effect was uniformly positive or negative (for interactions, the effect of the second design factor in the interaction) for all algorithms.", digits = 2, row.names = FALSE, escape = FALSE, booktabs = TRUE, col.names = gsub("(?<!vs)\\.", " ", names(tabanova), perl = TRUE)) |>
#   kable_styling(latex_options = c("scale_down"))
#tabanova <- tabanova[!grepl(":", tabanova$condition, fixed = T), ]
apa_table(tabanova, caption = "Effect size of design factors on predictive R2 of the different algorithms, and of the difference between algorithms. Interpretation indicates whether a main effect was uniformly positive or negative across all algorithms.", digits = 2, escape = FALSE, col.names = gsub("(?<!vs)\\.", " ", names(tabanova), perl = TRUE))#, longtable = TRUE)
# Interpretation for main effects indicates whether the effect was uniformly positive or negative or other, and for interactions, whether it is a spreading or crossover interaction.
```
```{r figr2, fig.cap="Predictive R2 for BRMA with horseshoe (HS) and LASSO prior, and RMA. Plots are sorted by largest performance difference between BRMA and RMA."}
knitr::include_graphics("r2.png")
```

@@

Cramer's V, an effect size for categorical variables, was used to examine the effect of design factors on sensitivity (Table \@ref(tab:tabcramerp), Figure \@ref(fig:figsensitivity)) and specificity (Table \@ref(tab:tabcramern), Figure \@ref(fig:figspecificity)).
We also computed this effect size for the difference between algorithms in the number of true positives by design factor.
Differences in sensitivity between the algorithms were near-zero for HS and LASSO.
The difference between the two BRMA algorithms and RMA were largest for the design factor effect size $\beta$, followed by the model and number of studies $k$.
For specificity, differences in sensitivity between HS and LASSO were largest for the number of noise moderators $M$, followed by the effect size $\beta$, number of studies $k$, and residual heterogeneity $\tau^2$.
The difference between the two BRMA algorithms and RMA were largest for the design factor number of studies $k$, followed by the model, the number of noise moderators $M$, and the effect size $\beta$.
Also note that the association between design factors and specificity was not monotonously positive or negative across algorithms.
Instead, some design factors had opposite effects for the two BRMA algorithms versus RMA.
For instance, a larger number of studies $k$ had a negative effect on specificity for the BRMA algorithms, but a positive effect for RMA - within the context that RMA had lower specificity on average.
Conversely, a greater number of noise moderators $M$ had a positive effect on specificity for BRMA, but a negative effect for RMA.
```{r tabcramerp, results= "asis", eval = TRUE}
tabselect <- readRDS("selected.RData")
rownames(tabselect) <- NULL
names(tabselect) <- gsub("(?<!vs)\\.", " ", names(tabselect), perl = TRUE)
apa_table(tabselect, caption = "Effect size (Cramer's V) of design factors, and of the difference between algorithms, on sensitivity (P).", digits = 2, escape = FALSE)#, longtable = TRUE)
# Interpretation for main effects indicates whether the effect was uniformly positive or negative or other, and for interactions, whether it is a spreading or crossover interaction.
```
```{r figsensitivity, fig.cap="Sensitivity by design factors for the HS (circle, solid line), LASSO(triangle, dotted line) and RMA (square, dashed line) algorithms."}
knitr::include_graphics("sensitivity.png")
```
```{r tabcramern, results= "asis", eval = TRUE}
tabselectn <- readRDS("notselected.RData")
rownames(tabselectn) <- NULL
names(tabselectn) = gsub("(?<!vs)\\.", " ", names(tabselectn), perl = TRUE)
apa_table(tabselectn, caption = "Effect size (Cramer's V) of design factors, and of the difference between algorithms, on specificity (N).", digits = 2, escape = FALSE)#, longtable = TRUE)
```
```{r figspecificity, fig.cap="Specificity by design factors for the HS (circle, solid line), LASSO(triangle, dotted line) and RMA (square, dashed line) algorithms."}
knitr::include_graphics("specificity.png")
```

@@

```{r exdata, eval = FALSE, echo = TRUE}
# Load relevant packages
library(pema)
library(mice)
library(rstan)
library(ggplot2)
# Select data to analyze
df <- bonapersona[ , c("yi", "vi", "mTimeLength",
                       "year", "model", "ageWeek")] 
# Multiple imputation for missing values
df <- mice(df)
```
```{r, eval = run_everything, echo = FALSE}
# Load relevant packages
library(pema)
library(mice)
library(rstan)
library(ggplot2)
# Select data to analyze
df <- bonapersona[ , c("yi", "vi", "mTimeLength",
                       "year", "model", "ageWeek")] 
# Multiple imputation for missing values
df <- mice(df)
saveRDS(df, "df_imputed.RData")
```
```{r, eval = !run_everything, echo = FALSE}
# Load relevant packages
library(pema)
library(mice)
library(rstan)
library(ggplot2)
df <- readRDS("df_imputed.RData")
```
<!-- Our simulation study shows good performance with default hyperparameters. -->
<!-- However, experienced users may want to customize the prior. -->
<!-- Visualizing the prior can be helpful in this process. -->
<!-- This is accomplished using the interactive application visualization application available through `shiny_prior()`. -->
<!-- The user can plot the prior distributions resulting from different sets of hyperparameters and compare them. -->
<!-- Increasing the values of the scale parameters (`scale_global` and `hs_scale_slab`) results in a more spread out prior, which applies less regularization. -->
<!-- Increasing the degrees of freedom (`df_global` and `df_slab`) results in thinner tails, which applies more regularization. -->
<!-- One advantage of the regularized horseshoe prior is its ability to define the prior in terms of the expected number of nonzero coefficients, using the argument `relevant_pars`. -->
<!-- This argument in turn influences the other hyperparameters. -->
<!-- We can also plot the implied prior on the effective number of nonzero coefficients given specific hyperparameters. -->
<!-- In this application, we have a total of 7 possible moderators, including the dummy variables for the categorical moderator "model". We can visualize the implied prior on the number of nonzero coefficients for different values of `relevant_pars`: p0 = 1; p0 = 3; p0 = 6; or p0 = NA, in which case no prior information is given and `scale_global`= 1. -->

@@

```{r, echo = FALSE, eval = FALSE}
summary(fit)
```
```{r tabsumex, eval = TRUE, echo = FALSE, results = "asis"}
tab <- summary(fit)$coefficients[1:8, c("mean", "sd", "2.5%", "50%", "97.5%", "n_eff", "Rhat")]
papaja::apa_table(tab, caption = "Summary of model parameters for the applied example.")
```
<!-- , see Figure \@ref(fig:figexres) -->
```{r exres, eval = FALSE, echo = FALSE}
fit_stan <- as.stan(fit)
plot(fit_stan, pars = rownames(coef(fit))[2:8])
```
```{r, eval = run_everything, echo = FALSE}
fit_stan <- as.stan(fit)
p <- plot(fit_stan, pars = rownames(coef(fit))[2:8])
ggsave("plot_es.png", p, device = "png", width = 4, height = 3, dpi = 300, scale = 1.5)
```
```{r figexres, eval = FALSE, echo = FALSE, fig.cap="Visual presentation of model parameters for the applied example."}
knitr::include_graphics("plot_es.png")
```
<!-- Note that there are many auxiliary parameters used for the regularized horseshoe prior. No information on these parameters is returned in the `brma` fitobject, however, once the fitobject is transformed into a Stan object for plotting, these parameters are included. -->

@@

As explained in the section on Bayesian estimation,
model convergence can be aided by increasing the amount of regularization of the prior, for example, by increasing some of the `df` parameters (see code below).
In this example, increasing both `df` and `df_slab` to 5 results in only 96 divergences, compared to the original 331.
This can be verified by running `summary(fit2)`.
In general, it is prudent to perform similar sensitivity analyses to determine how robust the results are to different priors.
For a visual inspection of the difference in posterior distributions,
use the function `plot_sensitivity()`, see Figure \@ref(fig:figsens).

```{r exmodel2, eval = FALSE, echo = TRUE}
fit2 <- brma(yi ~ ., data = df, vi = "vi",
             prior = c("df_global" = 5, "df" = 5),
             seed = 1)
plot_sensitivity(fit, fit2) + coord_cartesian(xlim = c(-3, 3))
```
```{r, echo = FALSE, eval = FALSE}
fit2 <- brma(yi ~ ., data = df, vi = "vi",
             prior = c("df_global" = 5, "df" = 5),
             seed = 1)
saveRDS(fit2, "fit2.RData")
p <- plot_sensitivity(fit, fit2) + coord_cartesian(xlim = c(-3, 3))
ggsave("plot_sensitivity.png", p, device = "png", units = "mm", width = 105, height = 60)
```
```{r figsens, echo = FALSE, eval = TRUE, fig.cap="A rudimentary visual check for sensitivity to different priors."}
knitr::include_graphics("plot_sensitivity.png")
```

>>>>

```{r tabr2eta, results= "asis", eval = TRUE}
tabanova <- readRDS("anova.RData")
tabanova <- tabanova[!grepl(":", tabanova$condition), ]
tabanova$Interpretation[!tabanova$Interpretation %in% c("positive", "negative", "other")] <- NA
#tabanova$condition <- gsub("alpha", "omega", tabanova$condition, fixed = T)
#tabanova$condition <- gsub("$mean{n}$", "$n$", tabanova$condition, fixed = T)
tabanova$Factor <- gsub("\\$(omega|beta|tau)", "\\$\\\\\\1", tabanova$Factor)
#tabanova$HS.vs..LASSO <- NULL
tabanova <- tabanova[, c(ncol(tabanova), 2:(ncol(tabanova)-1))]
# kbl(tabanova, caption = "Effect size of design factors and their two-way interactions on R2 the different algorithms, and of the difference between algorithms. The comparison between HS and LASSO was zero in the second decimal for all conditions. Interpretation indicates whether the effect was uniformly positive or negative (for interactions, the effect of the second design factor in the interaction) for all algorithms.", digits = 2, row.names = FALSE, escape = FALSE, booktabs = TRUE, col.names = gsub("(?<!vs)\\.", " ", names(tabanova), perl = TRUE)) |>
#   kable_styling(latex_options = c("scale_down"))
#tabanova <- tabanova[!grepl(":", tabanova$condition, fixed = T), ]
apa_table(tabanova, caption = "Effect size of design factors on predictive R2 of the different algorithms, and of the difference between algorithms. Interpretation indicates whether a main effect was uniformly positive or negative across all algorithms.", digits = 2, escape = FALSE, col.names = gsub("(?<!vs)\\.", " ", names(tabanova), perl = TRUE))#, longtable = TRUE)
# Interpretation for main effects indicates whether the effect was uniformly positive or negative or other, and for interactions, whether it is a spreading or crossover interaction.
```
```{r figr2, fig.cap="Predictive R2 for BRMA with horseshoe (HS) and LASSO prior, and RMA. Plots are sorted by largest performance difference between BRMA and RMA."}
knitr::include_graphics("r2.pdf")
```

@@

Cramer's V, an effect size for categorical variables, was used to examine the effect of design factors on sensitivity (Table \@ref(tab:tabcramerp), Figure \@ref(fig:figsensitivity)) and specificity (Table \@ref(tab:tabcramern), Figure \@ref(fig:figspecificity)).
We also computed this effect size for the difference between algorithms in the number of true positives by design factor.
Differences in sensitivity between the algorithms were near-zero for HS and LASSO.
The difference between the two BRMA algorithms and RMA were largest for the design factor effect size $\beta$, followed by the model and number of studies $k$.
For specificity, differences in sensitivity between HS and LASSO were largest for the number of noise moderators $M$, followed by the effect size $\beta$, number of studies $k$, and residual heterogeneity $\tau^2$.
The difference between the two BRMA algorithms and RMA were largest for the design factor number of studies $k$, followed by the model, the number of noise moderators $M$, and the effect size $\beta$.
Also note that the association between design factors and specificity was not monotonously positive or negative across algorithms.
Instead, some design factors had opposite effects for the two BRMA algorithms versus RMA.
For instance, a larger number of studies $k$ had a negative effect on specificity for the BRMA algorithms, but a positive effect for RMA - within the context that RMA had lower specificity on average.
Conversely, a greater number of noise moderators $M$ had a positive effect on specificity for BRMA, but a negative effect for RMA.
```{r tabcramerp, results= "asis", eval = TRUE}
tabselect <- readRDS("selected.RData")
rownames(tabselect) <- NULL
names(tabselect) <- gsub("(?<!vs)\\.", " ", names(tabselect), perl = TRUE)
apa_table(tabselect, caption = "Effect size (Cramer's V) of design factors, and of the difference between algorithms, on sensitivity (P).", digits = 2, escape = FALSE)#, longtable = TRUE)
# Interpretation for main effects indicates whether the effect was uniformly positive or negative or other, and for interactions, whether it is a spreading or crossover interaction.
```
```{r figsensitivity, fig.cap="Sensitivity by design factors for the HS (circle, solid line), LASSO(triangle, dotted line) and RMA (square, dashed line) algorithms."}
knitr::include_graphics("sensitivity.pdf")
```
```{r tabcramern, results= "asis", eval = TRUE}
tabselectn <- readRDS("notselected.RData")
rownames(tabselectn) <- NULL
names(tabselectn) = gsub("(?<!vs)\\.", " ", names(tabselectn), perl = TRUE)
apa_table(tabselectn, caption = "Effect size (Cramer's V) of design factors, and of the difference between algorithms, on specificity (N).", digits = 2, escape = FALSE)#, longtable = TRUE)
```
```{r figspecificity, fig.cap="Specificity by design factors for the HS (circle, solid line), LASSO(triangle, dotted line) and RMA (square, dashed line) algorithms."}
knitr::include_graphics("specificity.pdf")
```

@@

```{r exdata, eval = FALSE, echo = TRUE}
# Load relevant packages
library(pema)
library(mice)
library(rstan)
# Select data to analyze
df <- bonapersona[ , c("yi", "vi", "mTimeLength",
                       "year", "model", "ageWeek")] 
# Multiple imputation for missing values
df <- mice(df)
```
```{r, eval = run_everything, echo = FALSE}
# Load relevant packages
library(pema)
library(mice)
library(rstan)
# Select data to analyze
df <- bonapersona[ , c("yi", "vi", "mTimeLength",
                       "year", "model", "ageWeek")] 
# Multiple imputation for missing values
df <- mice(df)
saveRDS(df, "df_imputed.RData")
```
```{r, eval = !run_everything, echo = FALSE}
# Load relevant packages
library(pema)
library(mice)
library(rstan)
df <- readRDS("df_imputed.RData")
```
<!-- Our simulation study shows good performance with default hyperparameters. -->
<!-- However, experienced users may want to customize the prior. -->
<!-- Visualizing the prior can be helpful in this process. -->
<!-- This is accomplished using the interactive application visualization application available through `shiny_prior()`. -->
<!-- The user can plot the prior distributions resulting from different sets of hyperparameters and compare them. -->
<!-- Increasing the values of the scale parameters (`scale_global` and `hs_scale_slab`) results in a more spread out prior, which applies less regularization. -->
<!-- Increasing the degrees of freedom (`df_global` and `df_slab`) results in thinner tails, which applies more regularization. -->
<!-- One advantage of the regularized horseshoe prior is its ability to define the prior in terms of the expected number of nonzero coefficients, using the argument `relevant_pars`. -->
<!-- This argument in turn influences the other hyperparameters. -->
<!-- We can also plot the implied prior on the effective number of nonzero coefficients given specific hyperparameters. -->
<!-- In this application, we have a total of 7 possible moderators, including the dummy variables for the categorical moderator "model". We can visualize the implied prior on the number of nonzero coefficients for different values of `relevant_pars`: p0 = 1; p0 = 3; p0 = 6; or p0 = NA, in which case no prior information is given and `scale_global`= 1. -->

@@

```{r, echo = FALSE, eval = FALSE}
summary(fit)
```
```{r tabsumex, eval = TRUE, echo = FALSE, results = "asis"}
tab <- summary(fit)$coefficients[1:8, c("mean", "sd", "2.5%", "50%", "97.5%", "n_eff", "Rhat")]
papaja::apa_table(tab, caption = "Summary of model parameters for the applied example.")
```
<!-- , see Figure \@ref(fig:figexres) -->
```{r exres, eval = FALSE, echo = FALSE}
fit_stan <- as.stan(fit)
plot(fit_stan, pars = rownames(coef(fit))[2:8])
```
```{r, eval = run_everything, echo = FALSE}
fit_stan <- as.stan(fit)
p <- plot(fit_stan, pars = rownames(coef(fit))[2:8])
ggsave("plot_es.png", p, device = "png", width = 4, height = 3, dpi = 300, scale = 1.5)
ggsave("plot_es.pdf", p, device = "pdf", width = 4, height = 3, dpi = 300, scale = 1.5)
```
```{r figexres, eval = FALSE, echo = FALSE, fig.cap="Visual presentation of model parameters for the applied example."}
knitr::include_graphics("plot_es.pdf")
```
<!-- Note that there are many auxiliary parameters used for the regularized horseshoe prior. No information on these parameters is returned in the `brma` fitobject, however, once the fitobject is transformed into a Stan object for plotting, these parameters are included. -->

@@

As explained in the section on Bayesian estimation,
model convergence can be aided by increasing the amount of regularization of the prior, for example, by increasing some of the `df` parameters (see code below).
In this example, increasing both `df` and `df_slab` to 5 results in only 96 divergences, compared to the original 331.
This can be verified by running `summary(fit2)`.
In general, it is prudent to perform similar sensitivity analyses to determine how robust the results are to different priors.
For a visual inspection of the difference in posterior distributions,
use the function `plot_sensitivity()`.

```{r exmodel2, eval = FALSE, echo = TRUE}
fit2 <- brma(yi ~ ., data = df, vi = "vi",
             prior = c("df_global" = 5, "df" = 5),
             seed = 1)
plot_sensitivity(fit, fit2)
```
```{r, echo = FALSE, eval = FALSE}
fit2 <- brma(yi ~ ., data = df, vi = "vi",
             prior = c("df_global" = 5, "df" = 5),
             seed = 1)
saveRDS(fit2, "fit2.RData")
p <- plot_sensitivity(fit, fit2) + coord_cartesian(xlim = c(-3, 3))
ggsave("plot_sensitivity.png", p, device = "png", units = "mm", width = 105, height = 60)
ggsave("plot_sensitivity.pdf", p, device = "pdf", units = "mm", width = 105, height = 60)
```
```{r figsens, echo = FALSE, eval = FALSE, fig.cap="A rudimentary visual check for sensitivity to different priors."}
knitr::include_graphics("plot_sensitivity.pdf")
```
